\chapter*{I}

It was just before eight o'clock in the morning when Yakov Petrovich Golyadkin, a titular counselor, woke up after a long sleep, yawned, stretched, and finally opened his eyes wide. After a couple of minutes, however, he was lying motionless on his bed like a person who was not entirely sure whether he had woken up or was still asleep, whether what was happening around him was a waking reality or the disordered visions of a dream. Before long, however, Mr.~Golyadkin's senses began to take in their usual, everyday impressions more clearly and distinctly. The familiar dirty-green, sooty, dusty walls of his little room looked down on him, along with a red mahogany dresser, imitation red mahogany chairs, a table painted red, a Turkish sofa upholstered in reddish oilcloth with little green stars, and, finally, his clothes, removed helter-skelter the night before and thrown in a heap on the sofa. Finally, the grey autumn day, dreary and dirty, looked through the dull windows into the room so angrily and with such a sour grimace that Mr.~Golyadkin could no longer have any doubt that he found himself not in seventh heaven but in the city of Petersburg, the capital, on Shestilavochnaya Street, on the fourth floor of a very tall, solid building, in his own apartment. Having made this important discovery, Mr.~Golyadkin jerked his eyes shut, as if he wanted to get back the dream he had been having, if only for a minute. But after a minute, he jumped out of bed with a single leap, probably focusing, finally, on whatever idea his scattered thoughts had been circling until then, not finding their proper order. Having jumped out of bed, he immediately ran over to a small, round little mirror that stood on the dresser. Although the sleepy, short-sighted, and balding figure in the mirror was of an insignificant sort that at first glance did not draw any attention to itself in any way, nevertheless, its owner apparently remained completed satisfied with everything that he saw in the mirror. ``What a mess it would be,'' said Mr.~Golyadkin under his breath, ``what a mess it would be if something went wrong today, if it turned out, for example, that something was off---like if a pimple showed up somewhere or some other unpleasantness happened to me; however, things are not bad so far; so far everything is going fine.'' Delighted that everything was going fine, Mr.~Golyadkin put the mirror back where it had been and, despite the fact that he was barefoot and still wearing the same clothes in which he had the habit of nodding off at night, ran to the window and began with great interest to search for something in the building's courtyard, which the windows of his apartment faced. %golyadkin

