\part*{Part One}

\chapter*{I}

During a warm spell at the end of November, around nine o'clock in the morning, a train was approaching Petersburg at full speed on the Petersburg-Warsaw railway line. It was so foggy and damp that the rising sun could barely be seen; it was difficult to see anything more than ten paces away to the right and left of the tracks from the windows of the train. Some of the passengers were returning from abroad; but the third-class compartments were more crowded with people, all of them merchants and the lower classes from not very far away. Everyone was tired, as usual, everyone's eyes had become heavy during the night, everyone was freezing, their faces all pale yellow, the same color as the fog.

At sunrise, in one of the third-class cars, two passengers found themselves sitting across from one another, right by the window---both of them were young, both with barely any baggage, both far from elegantly dressed, both looked fairly unusual, and both of them, in the end, desired to strike up a conversation. If they had both known what precisely, in that moment, made the other unusual, then of course they would have been very surprised that the strange workings of chance had put them across from one another in a third-class car on the Petersburg-Warsaw train. One of them was short, about twenty-seven years old, with curly, almost black hair and small but fiery grey eyes. His nose was wide and flat, and he had high cheekbones; his thin lips incessantly formed themselves into an impertinent, mocking, even malicious smile; but he had a high brow, attractively formed, which made up for the crudely developed lower part of his face. That face was especially conspicuous for its deathly pallor, which gave the young man a worn-out look, despite his fairly strong build and something passionate, close to anguish, that was out of keeping with his crude, impudent smile and his sharp, self-satisfied gaze. He was dressed warmly in a broad, black lambskin coat lined in wool and did not feel chilly all night, while his neighbor was forced to endure all the delights of the dry Russian November night on his back, something for which he was obviously not prepared. He was wearing a fairly wide and thick sleeveless cloak with an enormous hood, exactly like the ones worn by travellers during winter somewhere far away, abroad, in Switzerland, for example, or Northern Italy, not counting on such destinations as Eydtkuhnen and Petersburg. But what was completely well-suited and satisfactory in Italy turned out to be not entirely fitting in Russia. The owner of the coat with the hood was a young man, also about twenty-six or twenty-seven years old, a little taller than average, with thick, very blond hair, sunken cheeks and a thin, pointed, almost completely white beard. His eyes were large, blue, and intense; there was something quiet but heavy in their gaze, something that could suggest a diagnosis of the falling sickness in the subject. The young man's face was pleasant, however, thin and cool, but colorless, and was by now frozen to the point of turning blue. He held in his hands a meager bundle wrapped in old, faded foulard, which seemed to contain all the possessions he had with him. He was wearing thick-soled shoes with gaiters---all very un-Russian. His black-haired neighbor in the wool-lined sheepskin coat took all of this in, in part because he had nothing better to do, and finally asked with an indelicate smirk, the kind that sometimes expresses, unceremoniously and carelessly, a person's pleasure at the misfortunes of others:

``Chilly?'' %parfyon

And he shrugged his shoulders.

``Very,'' his neighbor answered with extraordinary eagerness, ``and look, we're in the midst of a thaw. How bad would it be if it were really freezing? I didn't even think it could be this cold back home. I'm not used to it.'' %myshkin

``Coming from abroad, are you?'' %parfyon

``Yes, from Switzerland.'' %myshkin

``Phew! Quite a ways away!'' %parfyon

